---
title: "Testing Weighted NPMLE"
author: "Clifford Anderson-Bergman"
date: "7/12/2018"
output: html_document
---

This is a unit test for using weights for the NPMLE

```{r}
library(icenReg)
set.seed(123)
data(miceData)

group_fits = ic_np(cbind(l, u) ~ grp, data = miceData)
plot(group_fits)
```

Now we will fit "all" the data, but put zero weight on the ce group

```{r}
ce_only = rep(1, nrow(miceData))
ce_only[miceData$grp == 'ce'] = 0


ce_fit = ic_np(cbind(l,u) ~ 0, data = miceData, weights = ce_only)
plot(ce_fit)
```

We can see this the same fit as the "ce" group above. 

Now we will have the same data, but heavily weighted toward the "ge" group (rather than 100% weighted toward "ce" group)

```{r}
mostly_ce = rep(1, nrow(miceData))
mostly_ce[miceData$grp == "ce"] = 0.01

mostly_ce_fit = ic_np(cbind(l,u) ~ 0, data = miceData, weights = mostly_ce)
plot(mostly_ce_fit)
```

We will also compare with the results from `ic_sp`

```{r}
ic_sp_fit = ic_sp(cbind(l,u) ~ 0, data = miceData, weights = mostly_ce)
ic_sp_fit$llk - mostly_ce_fit$llk
# Difference in log likelihood is numerical error
```

Checking that fit is correct with random weights and regression formula

```{r}
rand_w = runif(nrow(miceData))

rand_w_np = ic_np(cbind(l, u) ~ grp, data = miceData, weights = rand_w)

ce_data = miceData[miceData$grp == "ce", ]
ce_w = rand_w[miceData$grp == "ce"]
rand_w_ce = ic_sp(cbind(l,u) ~ 0, data = ce_data, weights = ce_w)

rand_w_np$fitList$ce$llk - rand_w_ce$llk
```

Testing that weights work correctly when no formula but matrix is put in

```{r}
fit_np = ic_np(data = miceData[,c("l", "u")], weights = rand_w)
fit_sp = ic_sp(cbind(l, u) ~ 0, data = miceData, weights = rand_w)

fit_np$llk - fit_sp$llk
```